Ship-based lidar evaluation of Southern Ocean low clouds in the storm-resolving general circulation model ICON and the ERA5 and MERRA-2 reanalyses

### 2025-07-18 manuscript_agu_rev1.tex
Various

Some of the observatial data were likely used in reanalysis assimilation. The Macquarie Island station surface measurements and radiosonde profiles (not used in our analysis) were sent to the World Meteorological Organization Global Telecommunication System (GTS). The measurements on the RSV \textit{Aurora Australis} and HMNZS \textit{Wellington} were not used outside of research purposes. The AWS measurements but not lidar or radiosonde measurements on the RV \textit{Tangaroa} voyages were collected by the New Zealand MetService and communicated to the GTS. The ceilometer measurements on NBP1704 were not used outside of research purposes.

@@

When the data are subset by latitude (Fig.~\ref{fig:cloud-occurrence}b, c), we see that the low-latitude SO zone (40--55°S) displays a stronger peak of cloud occurrence near the surface than the high-latitude SO zone (between 55°S and the Antarctic coast), and this could be because higher latitudes have greater prevalence of weakly stable profiles (Fig.~\ref{fig:cyclone-stability}c, d), although more stable profiles populate regions south of 65°S close to the Antarctic coast. Cyclonic activity is also stronger in high-latitude SO, which is typically associated with shallow or deep convection rather than very stable stratification necessary for fog formation. The low- and high-latitude SO zones show similar biases in models as in the general case, but ERA5 does not overestimate the peak in the low-latitude SO zone (very low-level cloud occurrence is still strongly underestimated).

@@

Supercooled liquid clouds (liquid clouds under subzero temperature) commonly occur over the SO. In our analysis of LWP and IWP, we see that both phases are abundant. Because liquid water droplets are typically smaller and more numerous than ice crystals in cold clouds, they attenuate a greater amount of the lidar radiation. Clouds with a relatively modest optical thickness of 1.7 can attenuate the lidar signal for a detection at 2 km using an instrument with noise properties like the Vaisala CL31 (Section 2.4). While supercooled liquid clouds and their attenuation are accounted for by the lidar simulator, they can strongly attenuate the signal and cause artificially low values of cloud occurrence at higher altitudes. For example, we found that cloud occurrence at 1.5 km is underestimated in ICON and underlying clouds are overestimated. However, this can also mean that clouds at 1.5 km are present in the model but the signal is too attenuated by the lower clouds in the model but not in the observations, where the underlying clouds are not as pronounced.

@@

Subsetting by cyclonic activity and stability is done based on the ERA5 data. As we have shown, the reanalyses also suffer from biases in near-surface and upper-level quantities. Therefore, the subsetting is limited by the accuracy of the ERA5 pressure field, near-surface temperature, and temperature at 700~hPa. Near-surface ship observations are affected by the ship structures as well as the variable height above sea level at which the measurements are taken. The accuracy of radiosonde measurements in the first tens of meters from the surface is also likely affected by the ship environment, such as turbulence generated by ship structures and the ship exhaust. Vertical averaging of the radiosonde data can result in lower RH near saturation due to averaging of drier and moister layers together. For example, some of the RV \textit{Polarstern} radiosondes are available in vertical resolution of about 20--30 m. The satellite retrieval of LWP is affected by large biases, especially over high latitudes \cite{khanal2020}, which limits our comparison with the models.

@@

We suggest the following avenues for future research. Evaluation of ocean--atmosphere heat, moisture, and momentum fluxes with in-situ observations over the SO and comparison of GSRM simulations with large-eddy simulations in process-oriented studies; evaluation of the DYAMOND project simulations in a similar manner as performed here (for models which provide the necessary fields); and combining active satellite sensors such as the Cloud-Aerosol Lidar and Infrared Pathfinder Satellite Observations (CALIOP) on CALIPSO and Atmospheric Lidar [ATLID; \citeA{heliere2017}] on the Earth Clouds, Aerosols and Radiation Explorer [EarthCARE; \citeA{illingworth2015}] satellite with ground-based remote sensing could provide a more complete understanding of the cloud biases across the whole troposphere. Cloud phase could be analyzed in more detail using the CALIPSO data, as was done by \citeA{roh2020} in a cloud-resolving model, or using ground-based observations with a dual-polarization Mini Micro Pulse Lidar [MiniMPL; \citeA{spinhirne1993,campbell2002,flynn2007}] data available from the TAN1802 voyage. \citeA{guyot2022} and \citeA{whitehead2024} have developed a machine learning method for identifying cloud phase from ceilometer data, and this could be used with our ground-based lidar dataset to analyze the cloud phase. However, their method would require a careful calibration with reference data coming from this region.

>>>>

Some of the observational data were likely used in reanalysis assimilation. The Macquarie Island station surface measurements and radiosonde profiles (not used in our analysis) were sent to the World Meteorological Organization Global Telecommunication System (GTS). The measurements on the RSV \textit{Aurora Australis} and HMNZS \textit{Wellington} were not used outside of research purposes. The AWS measurements, but not lidar or radiosonde measurements on the RV \textit{Tangaroa} voyages, were collected by the New Zealand MetService and communicated to the GTS. The ceilometer measurements on NBP1704 were not used outside of research purposes.

@@

When the data are subset by latitude (Fig.~\ref{fig:cloud-occurrence}b, c), we see that the low-latitude SO zone (40--55°S) displays a stronger peak of cloud occurrence near the surface than the high-latitude SO zone (between 55°S and the Antarctic coast), and this could be because higher latitudes have a greater prevalence of weakly stable profiles (Fig.~\ref{fig:cyclone-stability}c, d), although more stable profiles populate regions south of 65°S close to the Antarctic coast. Cyclonic activity is also stronger in high-latitude SO, which is typically associated with shallow or deep convection rather than very stable stratification necessary for fog formation. The low- and high-latitude SO zones show similar biases in models as in the general case, but ERA5 does not overestimate the peak in the low-latitude SO zone (very low-level cloud occurrence is still strongly underestimated).

@@

Supercooled liquid clouds (liquid clouds under subzero temperature) commonly occur over the SO. In our analysis of LWP and IWP, we see that both phases are abundant. Because liquid water droplets are typically smaller and more numerous than ice crystals in cold clouds, they attenuate a greater amount of the lidar radiation. Clouds with a relatively modest optical thickness of 1.7 can attenuate the lidar signal for a detection at 2 km using an instrument with noise properties like the Vaisala CL31 (Section 2.4). While supercooled liquid clouds and their attenuation are accounted for by the lidar simulator, they can strongly attenuate the signal and cause artificially low values of cloud occurrence at higher altitudes. For example, we found that cloud occurrence at 1.5 km is underestimated in ICON and underlying clouds are overestimated. However, this can also mean that clouds at 1.5 km are present in the model but the signal is too attenuated by the lower clouds in the model, but not in the observations, where the underlying clouds are not as pronounced.

@@

Subsetting by cyclonic activity and stability is done based on the ERA5 data. As we have shown, the reanalyses also suffer from biases in near-surface and upper-level quantities. Therefore, the subsetting is limited by the accuracy of the ERA5 pressure field, near-surface temperature, and temperature at 700~hPa. Near-surface ship observations are affected by the ship structures as well as the variable height above sea level at which the measurements are taken. The accuracy of radiosonde measurements in the first tens of meters from the surface is also likely affected by the ship environment, such as turbulence generated by ship structures and the ship exhaust. Vertical averaging of the radiosonde data can result in lower RH near saturation due to averaging of drier and moister layers together. For example, some of the RV \textit{Polarstern} radiosondes are available in vertical resolution of about 20--30 m. As mentioned in Section~\ref{sec:rad-lwp-iwp}, the satellite retrieval of LWP and IWP is affected by large biases, especially over high latitudes, which limits our comparison with the models.

@@

We suggest the following avenues for future research. Evaluation of ocean--atmosphere heat, moisture, and momentum fluxes with in-situ observations over the SO and comparison of GSRM simulations with large-eddy simulations in process-oriented studies; evaluation of the DYAMOND project simulations in a similar manner as performed here (for models that provide the necessary fields); and combining active satellite sensors such as the Cloud-Aerosol Lidar and Infrared Pathfinder Satellite Observations (CALIOP) on CALIPSO and Atmospheric Lidar [ATLID; \citeA{heliere2017}] on the Earth Clouds, Aerosols and Radiation Explorer [EarthCARE; \citeA{illingworth2015}] satellite with ground-based remote sensing could provide a more complete understanding of the cloud biases across the whole troposphere. Cloud phase could be analyzed in more detail using the CALIPSO data, as was done by \citeA{roh2020} in a cloud-resolving model, or using ground-based observations with a dual-polarization Mini Micro Pulse Lidar [MiniMPL; \citeA{spinhirne1993,campbell2002,flynn2007}] data available from the TAN1802 voyage. \citeA{guyot2022} and \citeA{whitehead2024} have developed a machine learning method for identifying cloud phase from ceilometer data, and this could be used with our ground-based lidar dataset to analyze the cloud phase. However, their method would require a careful calibration with reference data coming from this region.
